PDBF Documentation

### 2015-08-05 Documentation.tex
small changes
removed unused folders
updated documentation and readme, added version file

\begin{document}
\maketitle

\section{Abstract}
PDBF documents are a hybrid format. They are a valid PDF and a valid HTML page at the same time. 
If you change the file extension to PDF and open it with an PDF viewer, you can see the static part of the document. If you change the file extension to HTML and open it with a Browser (currently Chrome/Firefox/Safari/IE 10 supported), you can see the dynamic part of the document. For example, if the static version contains an image of a chart which displays some data, then the dynamic version contains the actual raw data used to render the chart and renders the chart when opening the document. The advantage is that you can open the chart in an overlay view by clicking on the enlarge symbol at the upper left corner and then start to change parameters of the chart. For example you can remove filtering functions that are applied to the raw data or change confidence levels to other values an see the results of these changes directly in the chart. \\

\noindent PDBF files are created from LaTeX source code and a relational database. The raw data can either be a SQL statement string, a file with SQL statements, or contained in a database (currently PostgreSQL/MySQL/MariaDB supported). In the LaTeX code you can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements are supported) is created from the raw data with options and an SQL query. Read more in the [documentation](http://uds-datalab.github.io/PDBF/) (which is itself a PDBF document).\\

@@

\noindent This toolkit is licensed unter the MIT License.\\

\newpage
\noindent Thanks to the authors of:\\
phantomJS (\url{https://github.com/ariya/phantomjs})\\
Apache Commons IO (\url{http://commons.apache.org/proper/commons-io/})\\
Apache Commons Codec (\url{http://commons.apache.org/proper/commons-codec/})\\
google-gson (\url{https://github.com/google/gson})\\
postgreSQL JDBC Driver (\url{https://jdbc.postgresql.org/})\\
MariaDB JDBC Driver (\url{https://mariadb.com/kb/en/mariadb/about-the-mariadb-java-client/})\\
AlaSQL (\url{https://github.com/agershun/alasql})\\
C3 (\url{https://github.com/masayuki0812/c3})\\
D3 (\url{https://github.com/mbostock/d3})\\
Codemirror (\url{https://github.com/codemirror/codemirror})\\
google-diff-match-patch (\url{https://code.google.com/p/google-diff-match-patch/})\\
explorercanvas (\url{https://code.google.com/p/explorercanvas/})\\
DataTables (\url{https://github.com/DataTables/DataTables})\\
jQuery (\url{https://github.com/jquery/jquery})\\
jQuery-UI (\url{https://github.com/jquery/jquery-ui})\\
jStat (\url{https://github.com/jstat/jstat})\\
pivottable (\url{https://github.com/nicolaskruchten/pivottable})\\
PDF.js (\url{https://github.com/mozilla/pdf.js})\\
google-closure-compiler (\url{https://github.com/google/closure-compiler})\\
\newpage
\tableofcontents
\newpage

\section{Requirements}
These \LaTeX\relax packages are required for PDBF documents:\\
zref\\
xcolor\\
graphicx\\
xstring\\
xparse\\
geometry\\
array\\

\noindent Furthermore it is needed that your \LaTeX\relax document uses the geometry package to specify the page size.\\

\noindent \textbf{Warning:} There are problems with inputenc UTF8 package. If you really have to use the inputenc UTF8 package then you cant use non-ASCII characters inside queries.

@@

\noindent We extended AlaSQL with the following statistical sql functions:\\[5pt]
\verb|GRUBBS_FILTER(arr, alpha)|: \\
\indent Desc.: Filters an array of values with a \href{http://en.wikipedia.org/wiki/Grubbs'_test_for_outliers}{grubbs test}\\
\indent Returns the filtered array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\verb|MEAN(arr)|: \\
\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Arithmetic_mean}{arithmetic mean} of an array of values.\\
\indent Returns the arithmetic mean of the array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}\\[3pt]
%
\verb|STDDEV_SAMP(arr)|: \\
\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Standard_deviation}{ sample standard deviation} of an array of values.\\
\indent Returns the sample standard deviation of the array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}\\[3pt]
%
\verb|MARGIN_OF_ERROR(arr, alpha)|: \\
\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Margin_of_error}{margin of error} of an array of values.\\
\indent Returns the margin of error of the array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\verb|CONF_INT(arr, alpha)|: \\
\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Confidence_interval}{confidence interval} of an array.\\
\indent Returns the confidence interval of the array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\verb|T_TEST(arr1, arr2, alpha)|: \\
\indent Desc.: Performs a \href{http://en.wikipedia.org/wiki/Student's_t-test}{Student's t-test} on two arrays.\\
\indent Returns true if they are indistinguishable, and returns false otherwise.\\[3pt]
\indent \option
{arr1, arr2}
{Arrays on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\verb|WELCH_TEST(arr1, arr2, alpha)|:\\
\indent Desc.: Performs a \href{http://en.wikipedia.org/wiki/Welch's_t_test}{Welch's t-test} on two arrays.\\
\indent Returns true if they are indistinguishable, and returns false otherwise.\\[3pt]
\indent \option
{arr1, arr2}
{Arrays on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\section{Sql data sources}
\noindent Macros: \\[3pt]
\verb|\dbSQLText{sqlQueryString}| \\
\verb|\dbSQLFile{fileWithSqlQueries}| \\
\verb|\dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}| \\
\textbf{Note:} jdbcConnectionURL consist of jdbc followed by the name of the dbms (only postgresql and mysql are currently supported) followed by the url of the dbms followed by the database name (e.g. jdbc:postgresql://localhost:5432/postgres).\\
\textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output documents.

\newpage
\section{PDBF elements}
Currently there are four different PDBF elements: charts, pivot tables, multiplot charts, and sql statements.

\subsection{Chart}
\begin{figure}[h!]%
\hspace{-218pt}Examples:\\
    \centering
    \begin{minipage}{.48\textwidth}
    \chart[width=\textwidth, height=0.6\textwidth, chartType=line]{SELECT * FROM test;}
    \caption{Line Chart}
    \end{minipage}
    \hspace{11pt}
    \begin{minipage}{.48\textwidth}
    \chart[width=\textwidth, height=0.6\textwidth, chartType=bar]{SELECT * FROM test;}
    \caption{Bar Chart}
    \end{minipage} \\[8pt]

@@

\noindent Macros: \\[3pt]
\verb|\chart[options][queryForOverlay]{queryForPage}| \\
\textbf{Note:} if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay.\\
\textbf{Note:} The first column of the query result is used for the x-Axis, all other columns are used for y-Axis.\\[8pt] 
\noindent Options:\\[3pt]
%
\option
{name}
{Sets the name of this element. This is only useful for linking Elements. See DataText.}
{String}
{\texttt{undefined}}

@@

%\option %TODO: not yet implemented
%{logscale}
%{If set, the y axis uses log scale. Can also be redefined globally (\texttt{pdbfLogscale}).}
%{\texttt{true} or \texttt{false}}
%{\texttt{false}}
%
%\option %TODO: not yet implemented
%{overlap}
%{Unused.}
%{\texttt{TODO: ...}}
%{\texttt{TODO: ...}}
%
%\option %TODO: not yet implemented
%{legendpos}
%{Sets the position of the legend of the chart.}
%{\texttt{TODO: ...}}
%{\texttt{TODO: ...}}
%
%\option %TODO: not yet implemented
%{includeZero}
%{Set the minimum of the range of the y-axis to zero.}
%{\texttt{true} or \texttt{false}}
%{\texttt{false}}
%
%\option %TODO: not yet implemented
%{drawPoints}
%{Whether to show each point in line.}
%{\texttt{true} or \texttt{false}}
%{\texttt{false}}
%
%\option %TODO: not yet implemented
%{fillGraph}
%{Whether to fill the area below the graph.}
%{\texttt{true} or \texttt{false}}
%{\texttt{false}}
%
%\option %TODO: not yet implemented
%{showRangeSelector}
%{Whether to show a range selector for the x-axis below the chart.}
%{\texttt{true} or \texttt{false}}
%{\texttt{false}}

\newpage
\subsection{Multiplot Chart}
\begin{figure}[h!]%
    Example:\\
    \multiplotChart[width=\textwidth, height=0.5\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a > 2", "a <= 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}
    \caption{Multiplot Line Chart}
\end{figure}

@@

\section{F.A.Q.}
\begin{itemize}
\item Q.: The overlay is not on the right position.\\
A.: Most likely you use pages with different sizes in your document. This is currently not supported by PBDF 
\item Q.: I get \verb|Error: paperwidth value missing! Did you forgot to specify the| \\ \verb|papersize via the geometry package?| when compiling my document.\\
A.: You need to explicitly specify the papersize of your document via the geometry package (e.g. \verb|\usepackage[letterpaper]{geometry}|)
\item Q.: I get \verb|\unhbox \voidb@x \bgroup \let \unhbox \voidb@x \setbox \@tempboxa \hbox| \verb|{u\global \mathchardef \accent@spacefactor \spacefactor }\accent 127 u\egroup| \newline \verb|\spacefactor \accent@spacefactor| or a similar string in a error message.\\
A.: Chances are high that you use the \verb|\usepackage[utf8]{inputenc}| package or a similar package. These are not fully compatible with PDBF. If you really have to use the utf8 package then you cant use special characters inside options and queries of the PDBF package.
\end{itemize}

>>>>

\begin{document}
\maketitle
\noindent \textbf{For PDBF \input{VERSION.md}}

\section{Abstract}
PDBF documents are a hybrid format. They are a valid PDF and a valid HTML page at the same time. You can now optionally add an VirtualBox OVA file with a complete operating system to the PDBF document. Yes, this means that the resulting file is a valid PDF, HTML, and OVA file at the same time. If you change the file extension to PDF and open it with an PDF viewer, you can see the static part of the document. If you change the file extension to HTML and open it with a Browser (currently Chrome/Firefox/Safari/IE 10 supported), you can see the dynamic part of the document. And if an ova file is attached you can also change the file extension to OVA and install and run the attached operating system.\\
The difference between the PDF and the HTML version is that the PDF version contains static version of all PDBF elements, whereas the HTML version is dynamic. For example you can zoom into graphs, temporarly remove dataseries from the graph, inspect and change the underling query of the PDBF element and see the result of the change directly in the browser.\\
This approach works completely offline. No internet connection is required, neither at compile time, nor at viewing time.

\noindent PDBF files are created from LaTeX source code and a relational database. The raw data can either be a SQL statement string, a file with SQL statements, or contained in a database (currently PostgreSQL/MySQL/MariaDB supported). In the LaTeX code you can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements are supported) is created from the raw data with options and an SQL query. Read more in the \href{http://uds-datalab.github.io/PDBF/}{documentation}, which is itself is a PDBF document.\\

@@

\noindent A \href{https://infosys.uni-saarland.de/publications/p1972-dittrich.pdf}{demo paper} of our tool is appearing at \href{http://www.vldb.org/2015/}{VLDB 2015}.

\section{License}
\noindent This toolkit is licensed unter the MIT License. See LICENSE.md file.\\

\newpage
\tableofcontents
\newpage

\section{Thanks to}
phantomJS (\url{https://github.com/ariya/phantomjs})\\
Apache Commons IO (\url{http://commons.apache.org/proper/commons-io/})\\
Apache Commons Codec (\url{http://commons.apache.org/proper/commons-codec/})\\
google-gson (\url{https://github.com/google/gson})\\
postgreSQL JDBC Driver (\url{https://jdbc.postgresql.org/})\\
MariaDB JDBC Driver (\url{https://mariadb.com/kb/en/mariadb/about-the-mariadb-java-client/})\\
AlaSQL (\url{https://github.com/agershun/alasql})\\
C3 (\url{https://github.com/masayuki0812/c3})\\
D3 (\url{https://github.com/mbostock/d3})\\
Codemirror (\url{https://github.com/codemirror/codemirror})\\
google-diff-match-patch (\url{https://code.google.com/p/google-diff-match-patch/})\\
explorercanvas (\url{https://code.google.com/p/explorercanvas/})\\
DataTables (\url{https://github.com/DataTables/DataTables})\\
jQuery (\url{https://github.com/jquery/jquery})\\
jQuery-UI (\url{https://github.com/jquery/jquery-ui})\\
jStat (\url{https://github.com/jstat/jstat})\\
pivottable (\url{https://github.com/nicolaskruchten/pivottable})\\
PDF.js (\url{https://github.com/mozilla/pdf.js})\\
google-closure-compiler (\url{https://github.com/google/closure-compiler})\\

\section{How to use the PDBF compiler}
Right now there are only two modes: \\[4pt]
-- The standart mode which compiles a tex file into the hybrid PDBF format:\\
\verb|java -jar PDBF.jar LaTeX_file| \\[2pt]
-- The VM mode which attaches a VirtualBox .ova file to an existing PDBF:\\
\verb|java -jar PDBF.jar --vm PDBF_File.html VM_File.ova|

\section{Requirements}
These \LaTeX\relax packages are required for PDBF documents:
\begin{itemize}
\item \verb|zref|
\item \verb|xcolor|
\item \verb|graphicx|
\item \verb|xstring|
\item \verb|xparse|
\item \verb|geometry|
\item \verb|array|
\end{itemize}
\noindent Furthermore it is needed that your \LaTeX\relax document uses the geometry package to specify the page size.\\

\noindent \textbf{Warning:} There are problems with \textbf{inputenc UTF8} and \textbf{pgfpages}. If you really have to use the inputenc UTF8 package then you cant use non-ASCII characters inside queries. The pgfpages package is incompatible to this package (Its fixable but requires additional effort).

@@

\noindent We extended AlaSQL with the following statistical sql functions:\\[5pt]
\verb|GRUBBS_FILTER(arr, alpha)|: \\
\indent Desc.: Filters an array of values with a \href{http://en.wikipedia.org/wiki/Grubbs'_test_for_outliers}{grubbs test}\\
\indent Returns the filtered array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\verb|MEAN(arr)|: \\
\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Arithmetic_mean}{arithmetic mean} of an array of values.\\
\indent Returns the arithmetic mean of the array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}\\[3pt]
%
\verb|STDDEV_SAMP(arr)|: \\
\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Standard_deviation}{ sample standard deviation} of an array of values.\\
\indent Returns the sample standard deviation of the array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}\\[3pt]
%
\verb|MARGIN_OF_ERROR(arr, alpha)|: \\
\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Margin_of_error}{margin of error} of an array of values.\\
\indent Returns the margin of error of the array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\verb|CONF_INT(arr, alpha)|: \\
\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Confidence_interval}{confidence interval} of an array.\\
\indent Returns the confidence interval of the array.\\[3pt]
\indent \option
{arr}
{Array on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\verb|T_TEST(arr1, arr2, alpha)|: \\
\indent Desc.: Performs a \href{http://en.wikipedia.org/wiki/Student's_t-test}{Student's t-test} on two arrays.\\
\indent Returns true if they are indistinguishable, and returns false otherwise.\\[3pt]
\indent \option
{arr1, arr2}
{Arrays on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\verb|WELCH_TEST(arr1, arr2, alpha)|:\\
\indent Desc.: Performs a \href{http://en.wikipedia.org/wiki/Welch's_t_test}{Welch's t-test} on two arrays.\\
\indent Returns true if they are indistinguishable, and returns false otherwise.\\[3pt]
\indent \option
{arr1, arr2}
{Arrays on which the operation is performed.}
{JavaScript array}
{No default value. Value must always be set!}
\indent \option
{alpha}
{significance level}
{Number $>$ 0 AND Number $<$ 1}
{\texttt{0.05}}\\[3pt]
%
\section{SQL data sources}
\noindent Macros: \\[3pt]
\verb|\dbSQLText{sqlQueryString}| \\
\verb|\dbSQLFile{fileWithSQLQueries}| \\
\verb|\dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}| \\
\textbf{Note:} jdbcConnectionURL consist of jdbc followed by the name of the dbms (only postgresql and mysql are currently supported) followed by the url of the dbms followed by the database name (e.g. jdbc:postgresql://localhost:5432/postgres).\\
\textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output documents.\\[4pt]
\textbf{WARNING:} \verb|\dbSQLText| and \verb|\dbSQLFile| are currently horribly, even for medium sized (more than 100 tuples) amounts of data. Use \verb|\dbSQLJDBC| instead.

\section{PDBF elements}
\subsection{Chart}
\begin{figure}[h!]%
\hspace{-218pt}Examples:\\
    \centering
    \begin{minipage}{.48\textwidth}
    \chart[width=\textwidth, height=0.6\textwidth, chartType=line]{SELECT * FROM test;}
    \caption{Line Chart}
    \end{minipage}
    \hspace{11pt}
    \begin{minipage}{.48\textwidth}
    \chart[width=\textwidth, height=0.6\textwidth, chartType=bar]{SELECT * FROM test;}
    \caption{Bar Chart}
    \end{minipage} \\[8pt]

@@

\newpage
\noindent Macros: \\[3pt]
\verb|\chart[options][queryForOverlay]{queryForPage}| \\
\textbf{Note:} if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay.\\
\textbf{Note:} The first column of the query result is used for the x-Axis, all other columns are used for y-Axis.\\[8pt] 
\noindent Options:\\[3pt]
%
\option
{name}
{Sets the name of this element. This is only useful for linking Elements. See DataText.}
{String}
{\texttt{undefined}}

@@

\option
{includeZero}
{Set the minimum of the range of the y-axis to zero.}
{\texttt{true} or \texttt{false}}
{\texttt{false}}

\option
{drawPoints}
{Whether to show each point in line.}
{\texttt{true} or \texttt{false}}
{\texttt{false}}

\option
{fillGraph}
{Whether to fill the area below the graph.}
{\texttt{true} or \texttt{false}}
{\texttt{false}}

\option
{showRangeSelector}
{Whether to show a range selector for the x-axis below the chart.}
{\texttt{true} or \texttt{false}}
{\texttt{false} for in-page chart, \texttt{true} for overlay chart}

%\option %TODO: not yet implemented
%{logscale}
%{If set, the y axis uses log scale. Can also be redefined globally (\texttt{pdbfLogscale}).}
%{\texttt{true} or \texttt{false}}
%{\texttt{false}}
%
%\option %TODO: not yet implemented
%{overlap}
%{Unused.}
%{\texttt{TODO: ...}}
%{\texttt{TODO: ...}}
%
%\option %TODO: not yet implemented
%{legendpos}
%{Sets the position of the legend of the chart.}
%{\texttt{TODO: ...}}
%{\texttt{TODO: ...}}
%

\subsection{Multiplot Chart}
\begin{figure}[h!]%
    Example:\\
    \multiplotChart[width=\textwidth, height=0.5\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a > 2", "a <= 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}
    \caption{Multiplot Line Chart}
\end{figure}

@@

\section{F.A.Q.}
\begin{itemize}
\item Q.: The overlay is not on the right position!\\
A.: Most likely you use pages with different sizes in your document or you use the pgfpages package. This is currently not supported by PBDF 
\item Q.: I get \verb|Error: paperwidth value missing! Did you forgot to specify the| \\ \verb|papersize via the geometry package?| when compiling my document.\\
A.: You need to explicitly specify the papersize of your document via the geometry package (e.g. \verb|\usepackage[letterpaper]{geometry}|)
\item Q.: I get \verb|\unhbox \voidb@x \bgroup \let \unhbox \voidb@x \setbox \@tempboxa \hbox| \verb|{u\global \mathchardef \accent@spacefactor \spacefactor }\accent 127 u\egroup| \newline \verb|\spacefactor \accent@spacefactor| or a similar string in a error message.\\
A.: Chances are high that you use the \verb|\usepackage[utf8]{inputenc}| package or a similar package. These are not fully compatible with PDBF. If you really have to use the utf8 package then you cant use special characters inside options and queries of the PDBF package.
\end{itemize}
